Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-10-20 causal-bandit/nips_2016/additional_experiments.tex
updates experiment5
updates figures to remove type3 fonts
fits main paper back into 8 pages, makes description of additional experiments more concise

\begin{figure}
	\centering    
          \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=.45cm,
  thick,main node/.style={observed}, hidden/.style={empty},background rectangle/.style={fill=olive!45}]
%every node/.style={scale=0.6}
 %nodes
\node[main node](1){$X_{1}$};
\node[main node, right=of 1](2){$X_{2}$};
\node[hidden, right=of 2](3){$...$};
\node[main node, right=of 3](4){$X_{N}$};
\node[main node, below right=of 2](5){$Y$};
\node[main node, above right=of 2](6){$Z$};
 \path[every node/.style={font=\tiny}]
    (1) edge (5)
    	(2) edge (5)
    (4) edge (5)
    (6) edge (1) edge (2) edge (4);
\end{tikzpicture}
        \caption{Confounded graph}
        \label{fig:parallel_confounded} 
\end{figure}

@@

As in the parellel bandit case, we let $Y$ depend only on $X_1$, $P(Y|do(X_1)=1) = \frac{1}{2} + \epsilon$ amd $P(Y|do(X_1=0)) = \frac{1}{2}-\epsilon'$, where $\epsilon' = \epsilon\frac{P(X_1=1)}{P(X_1=0)}$. We vary $N_1$ from 2 to $N$ to vary $m$. The values for the CPD's have been chosen such that the reward distribution is independent of $m$ and so that we can analytically calculate $\eta*$. This allows us to just show the dependence on $m$, removing the noise associated with different models producing selecting values for $\eta*$ with the same $m$ (and thus same worst case performance) but different performance for a given reward distribution.

@@

In figure \ref{fig:simple_vs_T_misspecified} we exclude actions on $Z$ from the set of allowable actions to demonstrate how algorithm 1 can fail in the presence of a hidden confounding variable. Algorithm 1 incorrectly assumes $P(Y|do(X)) = P(Y|X)$ leading to biased estimates. To show the worst case outcome, where this bias results in Algorithm 1 asymptotically selecting a sub-optimal action, we need a slightly more complex CPD's.

\eq{
P(X_i = 1|Z = 0) &= 
\begin{cases} 
.166 & \text{ if } i \in \set{1,...6} \\
.2 & \text{ if } i = 6 \\
.7 & \text { otherwise} 
 \end{cases}\\
 P(X_i = 1|Z = 1) &= 
\begin{cases} 
.166 & \text{ if } i \in \set{1,...6} \\
.8 & \text{ if } i = 6 \\
.3 & \text { otherwise} 
 \end{cases}\\
}

$Y$ depends only on $X_6$ and $X_N$. 
\eq {
P(Y|\boldsymbol{X}) = \begin{cases} 
a & \text{ if } X_6=0,X_N = 0 \\
b & \text{ if } X_6=0,X_N = 1 \\
c & \text{ if } X_6=1,X_N = 1 \\
d & \text{ if } X_6=1,X_N = 0 \\
\end{cases}
}

\begin{figure}[H]
    \begin{subfigure}[t]{0.3\textwidth}
		\centering    
    		\includegraphics[width=\textwidth]{experiment4_20161012_1536.pdf}
    		\caption{Simple regret vs $m(\eta*)$ for fixed horizon $T=400$ and number of variables $N = 50$}
        \label{fig:simple_vs_m_general}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.3\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{experiment7_20161012_1147.pdf}
    		\caption{Simple regret vs horizon, $T$, with $N = 50$ and $m(\eta*)=3.1$ }
        \label{fig:simple_vs_T_general}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.3\textwidth}
    		\centering
    		\includegraphics[width=\textwidth]{experiment3_N50_s10000_20160519_1030.pdf}
    		\caption{Simple regret vs horizon, $T$, with $N = 50$, $m=xx$ with no actions setting $Z$}
    		\label{fig:simple_vs_T_misspecified}
    \end{subfigure}
    \caption{Experimental results on the confounded graph}
    \label{fig:experiments}
\end{figure}

>>>>

\begin{figure}[H]
	\centering    
          \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=.45cm,
  thick,main node/.style={observed}, hidden/.style={empty},background rectangle/.style={fill=olive!45}]
%every node/.style={scale=0.6}
 %nodes
\node[main node](1){$X_{1}$};
\node[main node, right=of 1](2){$X_{2}$};
\node[hidden, right=of 2](3){$...$};
\node[main node, right=of 3](4){$X_{N}$};
\node[main node, below right=of 2](5){$Y$};
\node[main node, above right=of 2](6){$Z$};
 \path[every node/.style={font=\tiny}]
    (1) edge (5)
    	(2) edge (5)
    (4) edge (5)
    (6) edge (1) edge (2) edge (4);
\end{tikzpicture}
        \caption{Confounded graph}
        \label{fig:parallel_confounded} 
\end{figure}

@@

As in the parallel bandit case, we let $Y$ depend only on $X_1$, $P(Y|do(X_1)=1) = \frac{1}{2} + \epsilon$ amd $P(Y|do(X_1=0)) = \frac{1}{2}-\epsilon'$, where $\epsilon' = \epsilon\frac{P(X_1=1)}{P(X_1=0)}$. We vary $N_1$ from 2 to $N$ to vary $m$. The values for the CPD's have been chosen such that the reward distribution is independent of $m$ and so that we can analytically calculate $\eta*$. This allows us to just show the dependence on $m$, removing the noise associated with different models producing selecting values for $\eta*$ with the same $m$ (and thus same worst case performance) but different performance for a given reward distribution.

@@

In figure \ref{fig:simple_vs_T_misspecified} we exclude actions on $Z$ from the set of allowable actions to demonstrate how algorithm 1 can fail in the presence of a hidden confounding variable. Algorithm 1 incorrectly assumes $P(Y|do(X)) = P(Y|X)$ leading to biased estimates. We let $P(Z) = .52$, $P(Y|\boldsymbol{X}) = X_7 \oplus X_N$ and $P(X|Z)$ be given by:

\eq{
P(X_i = 1|Z = 0) &= 
\begin{cases} 
.166 & \text{ if } i \in \set{1,..., 6} \\
.2 & \text{ if } i = 7 \\
.7 & \text { otherwise} 
 \end{cases}\\
 P(X_i = 1|Z = 1) &= 
\begin{cases} 
.166 & \text{ if } i \in \set{1,..., 6} \\
.8 & \text{ if } i = 7 \\
.3 & \text { otherwise} 
 \end{cases}\\
}

The first group of $X$ variables with $i \leq 6$ will be identified by the parallel bandit as the most unbalanced ones and played explicitly. All remaining variables are likely to be identified as balanced and estimated from observational estimates. This leads to the worst case outcome, where the bias in the estimates leads Algorithm 1 to asymptotically select a sub-optimal action.

\begin{figure}[H]
    \begin{subfigure}[t]{0.3\textwidth}
		\centering    
    		\includegraphics[width=\textwidth]{experiment4_20161020_1254.pdf}
    		\caption{Simple regret vs $m(\eta*)$ for fixed horizon $T=400$ and number of variables $N = 50$}
        \label{fig:simple_vs_m_general}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.3\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{experiment7_20161020_1257.pdf}
    		\caption{Simple regret vs horizon, $T$, with $N = 50$ and $m(\eta*)=3.1$ }
        \label{fig:simple_vs_T_general}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.3\textwidth}
    		\centering
    		\includegraphics[width=\textwidth]{experiment5_20161020_2027.pdf}
    		\caption{Simple regret vs horizon, $T$, with $N = 20$, $m=4.3$ with no actions setting $Z$}
    		\label{fig:simple_vs_T_misspecified}
    \end{subfigure}
    \caption{Experimental results on the confounded graph}
    \label{fig:experiments}
\end{figure}
